Hierarchical Reverberation Mapping

### 2013-09-29 Paper/ms.tex
Various edits

There are many subtleties involved in reverberation mapping, that we will ignore
for the purposes of this letter. We note them here for completeness. Firstly,
the mean lag $\bar{\tau} = \int \Psi(\tau)\, d\tau$, where $\Psi(\tau)$
is the transfer function, is not equal to $c$ times the mean radius of the BLR
distribution. Secondly, the mean lag $\bar{\tau}$ is also not what is measured
by the common
cross-correlation technique, except in the case of very narrow
transfer functions. Direct physical modelling of the BRL resolves these
issues \citep{pancoast, arp151}.

@@

If the continuum light curve is described by a function $y(t)$, then the line
light curve $l(t)$ is given by
\begin{eqnarray}
l(t) &=& A \int_\tau \Psi(\tau)\left[y(t - \tau) + C\right] \, d\tau
\end{eqnarray}
where $\Psi(\tau)$ is the transfer function (assumed to be normalised),
and $A$ and $C$ are response
coefficients. The idea of reverberation mapping is to use noisy measurements
of $y(t)$ and $l(t)$ to infer the transfer function $\Psi(\tau)$ or a summary
of it such as the mean lag $\bar{\tau} = \int \tau\Psi(\tau) \, d\tau$.
Throughout this letter we consider $y(t)$ and $l(t)$ as fluxes, not magnitudes.

@@

In a hierarchical model, the prior for the parameters $\{\theta_i\}$ is created
by introducing hyperparameters $\alpha$ such that the joint prior for $\alpha$
and the $\{\theta_i\}$ is:
\begin{eqnarray}
p(\alpha, \theta_i) &=& p(\alpha)\prod_{i=1}^N p(\theta_i | \alpha)
\end{eqnarray}
The posterior distribution for
$\btheta = \{\theta_1, ..., \theta_N\}$ and the hyperparameters $\alpha$
given the data 
$\bx = \{x_1, ..., x_N\}$
should actually be
\begin{eqnarray}
p(\alpha, \btheta | \bx) &\propto&
p(\alpha)p(\btheta|\alpha)p(\bx | \btheta, \alpha)\\
&=& p(\alpha)\prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i)
\end{eqnarray}
The marginal posterior distribution for the hyperparameters $\alpha$ is
\begin{eqnarray}
p(\alpha | \bx) &=&
\int p(\alpha, \btheta|\bx) \, d\btheta\\
&\propto& p(\alpha)\int \prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i) d^N\theta\\
&\propto& p(\alpha) \prod_{i=1}^N \int f(\theta_i|\alpha)p(x_i | \theta_i) d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \int \frac{f(\theta_i|\alpha)}{\pi(\theta)}p(x_i | \theta_i) \pi(\theta_i)d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \mathds{E}\left[\frac{f(\theta_i|\alpha)}{\pi(\theta)}\right]
\end{eqnarray}
where the expectation is taken with respect to the individual object posterior
of Equation~\ref{eq:individual}. This result enables us to reconstruct the
posterior distribution for the hyperparameters even though the individual object
inferences were made without the hierarchical structure in the prior.

\section{The Single Object Model}
The posterior distribution for the mean lag $\bar{\tau}$ of a single object
$i$ is obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}s
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag $\bar{\tau} = \frac{1}{2}(b-a)$.

The prior for the underlying time variation of the continuum emission is
a continuous autoregressive process of order 1, or a CAR(1) model. These models
have been studied extensively for AGN variability
\citep[e.g.][]{2009ApJ...698..895K, 2011ApJ...735...80Z, 2013ApJ...765..106Z}.
Note that inferring $\bar{\tau}$ from the data requires that we
marginalise over an infinite number of nuisance parameters describing the
behaviour of $y(t)$ at unobserved times \citep{pancoast}.

This marginalisation can be done
either analytically or inside MCMC. We use the latter approach. We discretise
time using ten time bins per day, and have a discrete continuum light curve
$\mathbf{y} = \{y_1, ..., y_n\}$
\begin{eqnarray}
y_i = \mu + a\left(y_{i-1} - \mu\right) + \beta n_i
\end{eqnarray}
for $i \geq 2$, where the prior for the $\{n_i\}$ is independent standard
normal distributions. This is the discrete AR(1) model from time series theory.

@@

For the MCMC we used STAN \citep{nuts} and DNest \citep{dnest}.

Note the single-object model is already an improvement over standard
cross-correlation techniques.

\section{Demonstration on Simulated Data}

>>>>

There are many subtleties involved in reverberation mapping, that we will ignore
for the purposes of this letter. We note them here for completeness. Firstly,
the mean lag $\bar{\tau} = \int \tau\Psi(\tau)\, d\tau$, where $\Psi(\tau)$
is the normalised transfer function, is not equal to $c$ times the mean radius of the BLR
distribution. Secondly, the mean lag $\bar{\tau}$ is also not what is measured
by the common
cross-correlation technique, except in the case of very narrow
transfer functions. Direct physical modelling of the BRL resolves these
issues \citep{pancoast, arp151}.

@@

If the continuum light curve is described by a function $y(t)$, then the line
light curve $l(t)$ is given by
\begin{eqnarray}
l(t) &=& A \int_\tau \Psi(\tau)\left[y(t - \tau) + C\right] \, d\tau
\end{eqnarray}
where $\Psi(\tau)$ is the transfer function (assumed to be normalised),
and $A$ and $C$ are response
coefficients. The idea of reverberation mapping is to use noisy measurements
of $y(t)$ and $l(t)$ to infer the transfer function $\Psi(\tau)$ or a summary
of it such as the mean lag $\bar{\tau} = \int \tau\Psi(\tau) \, d\tau$.
Throughout this letter we consider $y(t)$ and $l(t)$ in flux
units, as opposed to magnitudes.

@@

In a hierarchical model, the prior for the parameters $\{\theta_i\}$ is created
by introducing hyperparameters $\alpha$ such that the joint prior for $\alpha$
and the $\{\theta_i\}$ is:
\begin{eqnarray}
p(\alpha, \theta_i) &=& p(\alpha)\prod_{i=1}^N p(\theta_i | \alpha)
\end{eqnarray}
The posterior distribution for
$\btheta = \{\theta_1, ..., \theta_N\}$ and the hyperparameters $\alpha$
given the data 
$\bx = \{x_1, ..., x_N\}$
should actually be
\begin{eqnarray}
p(\alpha, \btheta | \bx) &\propto&
p(\alpha)p(\btheta|\alpha)p(\bx | \btheta, \alpha)\\
&=& p(\alpha)\prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i)
\end{eqnarray}
The marginal posterior distribution for the hyperparameters $\alpha$ is
\begin{eqnarray}
p(\alpha | \bx) &=&
\int p(\alpha, \btheta|\bx) \, d\btheta\\
&\propto& p(\alpha)\int \prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i) d^N\theta\\
&\propto& p(\alpha) \prod_{i=1}^N \int f(\theta_i|\alpha)p(x_i | \theta_i) d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \int \frac{f(\theta_i|\alpha)}{\pi(\theta)}p(x_i | \theta_i) \pi(\theta_i)d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \mathds{E}\left[\frac{f(\theta_i|\alpha)}{\pi(\theta)}\right]
\end{eqnarray}
where the expectation is taken with respect to the individual object posterior
of Equation~\ref{eq:individual}, and thus can be estimated using posterior
samples from the posterior distributions for the individual objects.
This result enables us to reconstruct the
posterior distribution for the hyperparameters even though the individual object
inferences were made without the hierarchical structure in the prior. This is
essentially an importance sampling approximation to the hierarchical model.

\section{The Single Object Model}
The posterior distribution for the mean lag $\bar{\tau}$ of a single object
$i$ may be obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}s
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag:
\begin{eqnarray}
\bar{\tau} &=& \int \\
&=& \frac{1}{2}(b-a)
\end{eqnarray}

Note that inferring $\bar{\tau}$ from the data requires that we
marginalise over an infinite number of nuisance parameters describing the
behaviour of $y(t)$ at unobserved times \citep{pancoast}.
The prior for the underlying time variation of the continuum emission is
a continuous autoregressive process of order 1, or a CAR(1) model. These models
have been studied extensively for AGN variability
\citep[e.g.][]{2009ApJ...698..895K, 2011ApJ...735...80Z, 2013ApJ...765..106Z}.
This marginalisation can be done
either analytically or inside MCMC. We used the latter approach for simplicity.
We discretised
time using ten time bins per day, and have a discrete continuum light curve
$\mathbf{y} = \{y_1, ..., y_n\}$
\begin{eqnarray}
y_i = \mu + a\left(y_{i-1} - \mu\right) + \beta n_i
\end{eqnarray}
for $i \geq 2$, where the prior for the $\{n_i\}$ is independent standard
normal distributions. This is the discrete AR(1) model from time series theory.
$\mu$ describes the mean level of the continuum light curve, $\beta$ controls
the size of the short-term fluctuations, and $a$ controls the correlation
timescale.

@@

To implement the MCMC we implemented this model in the STAN sampler \citep{nuts}
and Diffusive Nested Sampling \citep{dnest}. Due to the large number of
parameters and some very strong correlations in the posterior distribution,
we found that Diffusive Nested Sampling was more effective than STAN.
Note the single-object model is already an improvement over standard
cross-correlation techniques and is similar to the approach used by
\citep{2011ApJ...735...80Z}.

\section{Demonstration on Simulated Data}
To test our hierarchical model, and compare it to the stacked cross-correlation
function, we simulated data from 100 AGN.

The cross-correlation function for a single object is
